---
title: "HIWI2020 test"
author: "Emmanuel Adeleke"
date: "9/28/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the required libraries

```{r message = F}
library(raster)
library(ncdf4)
library(dplyr)
```


## Open NCDFile as raster object

```{r warning=FALSE}

prcp <- brick("prcp-195101-grd-scaled.nc", var = "prcp")
prcp[prcp == NaN] <- NA

```

## Plot raster map e.g. day 1

```{r}
plot(prcp$X1951.01.01)
```

Prepare basin boundary information from the BDY file
=============================

```{r}
# read BDY file as text file
bound <- read.delim("02387000.BDY")

# create empty matrix to populate later in a for loop
out <- matrix(ncol=2, nrow=length(bound$X023870....621))

for (i in 1:length(bound$X023870....621)){
  out[i,] <- strsplit(bound$X023870....621[i], split = " ", fixed = T)[[1]][c(2,4)]
}

# convert matrix to dataframe
df_bound <- as.data.frame(out)

# convert string columns to numeric type and rename columns 
df_bound$V1 <- as.numeric(df_bound$V1)
df_bound$V2 <- as.numeric(df_bound$V2)
names(df_bound) <- c("Longitude", "Latitude")


```

Here, we convert the boundary dataframe to a SpatialPolygons R object
=======================================================================

```{r warning =FALSE}

# convert coordinates to SpatialPolygon object 
poly <- Polygon(df_bound[,c(1,2)], hole = T)
poly_list <- Polygons(list(poly), ID = "polygon1")
boundPolygon <- SpatialPolygons(list(poly_list), proj4string = CRS('+proj=longlat +datum=WGS84 +no_defs'))
boundPolygon
```

```{r basin}
# plot basin polygon 
plot(boundPolygon)

```

Now we overlap the basin on the raster map and extract relevant information
============================================================================
Generated 'results' is a list containing raster cell number(ID),
daily precipitation estimate in that cell, fraction(%) of cell that is covered by polygon,
and the coordinates of the cells

```{r warning =FALSE}
# create empty list to populate in a for loop
results <- list()

for (i in 1:nlayers(prcp)){
  
  # overlay basin on raster and extract daily precip data and percentage of overlapping region
  results[[i]] <- extract(prcp[[i]], boundPolygon, weights = T, cellnumbers=T)
  
  # extract coordinates of polygons and bind to data
  results[[i]] <- cbind(results[[i]][[1]], coordinates(prcp[[i]])[results[[i]][[1]][,1],])
  
}
```


Prepare result dataframe for export
=====================================

```{r}
# convert list 'results' to a dataframe
finalDF<- do.call("rbind", results) %>% as.data.frame()
finalDF$weight <- finalDF$weight*100

# rename columns
names(finalDF) <- c("cell_ID", "daily_precip", "cell_percentage", "Longitude", "Latitude")

```

We can also calculate weighted average for each cell that overlaps with the polygon
====================================================================================

```{r}
finalDF$weighted_avg <- (finalDF$daily_precip*finalDF$cell_percentage)/finalDF$cell_percentage
head(finalDF)
```

Finally, we calculate weighted daily average precipitation
============================================================

```{r}
# Calculate daily average precipitation (pavg) 

# create empty vector to populate in a for loop
pavg <- c()

for (i in 1:length(results)){
  pavg[i] <- sum(results[[i]][,2] * (results[[i]][,3]*100)) / sum(results[[i]][,3]*100)
  
}

# convert data to a dataframe 
pavg <- as.data.frame(pavg)
head(pavg)

```

# Export results in .csv files

```{r}
write.csv(finalDF, "results/weightedavgprecip.csv")
write.csv(pavg, "results/weightedavgdailyprecip.csv")

```

